# Solution to Chapter 2 Exercises

```{r}
#| message: false
#| warning: false
library(fpp3)
```


## Exercise 1 {.unnumbered}

Explore the following four time series: `Bricks` from `aus_production`, `Lynx` from `pelt`, `Close` from `gafa_stock`, `Demand` from `vic_elec`.

    a. What is the time interval of each series?
    b. Use autoplot() to produce a time plot of each series. For the last plot, modify the axis labels and title.
    
    
### (a) {.unnumbered}

```{r}
tsibble::interval(aus_production)
tsibble::interval(pelt)
```


```{r}
gafa_stock %>% select(Close)
```

```{r}
tsibble::interval(gafa_stock)
```


`gafa_stock` data has irregular time interval.

```{r}
tsibble::interval(vic_elec)
```

`vic_elec` data has a time interval of 30 mins.


### (b) {.unnumbered}

```{r}
aus_production %>% autoplot(Bricks)
pelt %>% autoplot(Lynx)
gafa_stock %>% autoplot(Close) + labs(x = NULL)
vic_elec %>% 
  autoplot(Demand) +
  labs(title = "Time series of Demand from vic_elec", x = NULL)
```


## Exercise 2 

Use `filter()` to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.

```{r}
gafa_stock %>% 
  group_by(Symbol) %>%
  filter(Close == max(Close)) %>% 
  select(Symbol, Date, Close) %>% ungroup()
```


## Exercise 3

Download the file `tute1.csv` from the [book website](https://bit.ly/fpptute1), open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

    (a) Read the data and create time series
    (b) Construct ts plot for each series

```{r}
#| message: false
tute1 <- readr::read_csv(here::here("data/tute1.csv"))
head(tute1)
```

```{r}
tute1 %>% 
  mutate(Quarter = yearquarter(Quarter)) %>% 
  as_tsibble(index = Quarter) -> ts_tute

ts_tute
```


```{r}
ts_tute %>% 
  pivot_longer(-Quarter) %>% autoplot(value)
```

```{r}
ts_tute %>% 
  pivot_longer(-Quarter) %>% 
  ggplot(aes(x = Quarter, y = value, color = name)) +
  geom_line()
```

```{r}
ts_tute %>% 
  pivot_longer(-Quarter) %>% 
  ggplot(aes(x = Quarter, y = value, color = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```

## Exercise 4

The `USgas` package contains data on the demand for natural gas in the US.

(a) Create a tsibble from us_total with year as the index and state as the key.
(b) Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).

```{r}
data(us_total, package = "USgas")

head(us_total)
```

```{r}
us_total %>% 
  as_tsibble(index = year, key = state) -> ts_us_total

ts_us_total
```

```{r}
ts_us_total %>% 
  filter(state %in% c("Maine", "Vermont", "New Hampshire", "Massachusetts", "Connecticut",
                      "Rhode Island")) %>% 
  autoplot(y) +
  scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale()))
```

## Exercise 5

(a) Download `tourism.xlsx` from the [book website](https://bit.ly/fpptourism) and read it into R using `readxl::read_excel()`.
(b) Create a `tsibble` which is identical to the tourism tsibble from the tsibble package.
(c) Find what combination of Region and Purpose had the maximum number of overnight trips on average.
(d) Create a new `tsibble` which combines the Purposes and Regions, and just has total trips by State.

```{r}
tourism_xl <- readxl::read_excel("data/tourism.xlsx")
head(tourism_xl)
```

```{r}
tourism
```



```{r}
tourism_xl %>% 
  mutate(Quarter = yearquarter(Quarter)) %>% 
  as_tsibble(index = Quarter, key = c(Region, State, Purpose)) -> tourism_ts

tourism_ts
```

```{r}
tourism_ts %>% 
  group_by(Region, Purpose) %>% 
  summarise(trips = sum(Trips), .groups = "drop") %>%
  slice_max(trips, n = 1)
```

```{r}
tourism_ts %>% 
  group_by(State) %>% 
  summarise(trips = sum(Trips), .groups = "drop") -> trips_by_state

trips_by_state
```


## Exercise 6

The `aus_arrivals` data set comprises quarterly international arrivals to Australia from Japan, New Zealand, UK and the US.

(a) Use `autoplot()`, `gg_season()` and `gg_subseries()` to compare the differences between the arrivals from these four countries.
(b) Can you identify any unusual observations?

```{r}
aus_arrivals
```

```{r}
aus_arrivals %>% 
  autoplot(Arrivals) +
  scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale()))
```

```{r}
aus_arrivals %>% 
  gg_season(Arrivals) +
  scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale()))
```

```{r}
aus_arrivals %>% gg_subseries(Arrivals)
```

